---
title: "Feasibility Analysis in R"
output: html_notebook
---

Run feasibility analysis for ASH 2022 Abstract submission.
Potential study topics:
a. COVID19 Vaccine >> comparing thrombotic events across vaccine types
b. COVID19 Vaccine utilization in hematological malignancies compared to control group
c. Oxbryta >> vaso-occlusive events compared with conventional treatment
d. Selinexor >> Thrombocytopenia and leukopenia compared to conventional treatment
e. Dorzalex adherence

```{r config}
workdir <- "~/komodo_research/projects/ash_2022"
library(tidyverse)
library(glue)
library(clipr)
library(tictoc)
library(DBI)
library(googlesheets4)
source("~/komodo_research/library/R/utils.R")
source("~/komodo_research/library/R/sql_helper.R")

read_query <- function(sql, conn = con_odbc) {
  df <- DBI::dbGetQuery(conn, sql)
  colnames(df) <- tolower(colnames(df))
  df
}

read_table <- function(..., conn = con_odbc) {
  # tyipical input of ... expects:
  # database, schema, table
  tbl(conn, DBI::Id(...))
}

execute_sql <- function(sql, conn = con_odbc, print = FALSE) {
  if (print) cat(sql)
  DBI::dbExecute(conn, sql)
  return(invisible(NULL))
}

# 
# time_it_fun <- function(before, options) {
#   require(tictoc)
#   if (before) {
#     # record the current time before each chunk
#     # now <<- Sys.time()\
#     tic(options$label)
#   } else {
#     # calculate the time difference after a chunk
#     # res <- difftime(Sys.time(), now)
#     toc()
#     # return a character string to show the time
#     # paste("Time for the chunk", options$label, "to run:", res)
#   }
# }
# 
# 
# # timeit
# knitr::knit_hooks$set(time_it = local({time_it_fun}))
# knitr::opts_chunk$set(time_it = TRUE)
```

```{r connection}
# create ODBC connection
con_odbc <- DBI::dbConnect(
  odbc::odbc(), "snowflake", 
  role = "ANALYST", 
  warehouse = "LARGE_WH"
)

execute_sql("USE DATABASE SANDBOX_KOMODO;")
execute_sql("USE SCHEMA AYWEI;")
execute_sql("USE ROLE ANALYST;")
execute_sql("USE WAREHOUSE XLARGE_WH;")
```

```{r gsconnection}
# google sheet auth
gs4_auth(
  email = gargle::gargle_oauth_email(),
  path = NULL,
  scopes = "https://www.googleapis.com/auth/spreadsheets",
  cache = gargle::gargle_oauth_cache(),
  use_oob = gargle::gargle_oob_default(),
  token = gargle::token_fetch("https://www.googleapis.com/auth/spreadsheets")
)

ss = "1Vs8wMBUndpKMF1jx_PVWZskhREAGYaJUudteNMNvviM"
code_list_ss = "1lwGATVIUxa1RMtcfUmpZRoJLEcDadBboyTmTthN2U20"
```

```{r code_list}
# create a code list for scd, voc and oxbryta
voc_diag_sql = get_code(c('D570%', 'D5721%', 'D5741%', 'D5743%', 'D5745%', 'D5781%'), method = 'pattern', type = 'diag')
scd_diag_sql = get_code(c('D570%', 'D571%', 'D572%', 'D574%', 'D578%'), method = 'pattern', type = 'diag')
oxbryta_ndc_sql = get_code(c('oxbryta'), method = 'name', type = 'ndc')
print(voc_diag <- read_query(voc_diag_sql))
print(scd_diag <- read_query(scd_diag_sql))
print(oxbryta_ndc <- read_query(oxbryta_ndc_sql))
sheet_write(voc_diag, ss = ss, sheet = 'voc_diag')
sheet_write(scd_diag, ss = ss, sheet = 'scd_diag')
sheet_write(oxbryta_ndc, ss = ss, sheet = 'oxbryta_ndc')
```

```{r code_list2}
# create code_list
code_list <- 
  bind_rows(
    voc_diag %>% select(code, code_description) %>% mutate(codetype = "ICD10CM", event = "voc"),
    scd_diag %>% select(code, code_description) %>% mutate(codetype = "ICD10CM", event = "scd"),
    oxbryta_ndc %>% select(ndc, cui_l1_name) 
      %>% rename(code = ndc, code_description = cui_l1_name) 
      %>% mutate(codetype = "NDC", event = "oxbryta")
  )
write_csv(code_list, file = file.path(workdir, "code_list.csv"))
```

```{r upload_code_list}
bulk_upload(
  conn = con_odbc,
  dir = workdir,
  filename = "code_list.csv", 
  stage_name = "stage_kr",
  table_full_name = "SANDBOX_KOMODO.AYWEI.SCD_CODE_LIST",
  create_stage = FALSE,
  database = 'SANDBOX_KOMODO',
  role = 'ANALYST',
  warehouse = 'LARGE_WH',
  save_sql = TRUE,
  run_sql = TRUE,
  snowsql_profile = 'ywei'
)
```

```{r project_spec}
## Table reference
rx_version = '20220613'
mx_version = '20220605'
code_version = '20220606'
query_start_date = '2019-01-01'
query_end_date = '2022-12-31'
code_list_table = 'scd_code_list'
## Output table names
prefix = 'SCD'

```

```{r step_1_get_ce}
# build CE table for all patients
sql_split_fun = get_sql_split_fun(grace_period = grace_period)
execute_sql(sql_split_fun)
tic("CE")
sql = glue("  
    create or replace table ce_{mx_version} as
    {get_ce(mx_version = mx_version, grace_period = 45)}
    ;
")
execute_sql(sql)
toc()
```

```{r step_2_get_query}
# build encounter tables
tic("Mx Encounter")
sql_mx = glue("
    create or replace table {prefix}_mx_encounter as
    {query_event_mx(
      code_list_table = code_list_table,
      mx_version = mx_version,
      start_date = query_start_date, end_date = query_end_date
      )
    }
    ;
")
execute_sql(sql_mx)
toc()
# 
# tic("Rx Encounter")
# sql_rx = glue("  
#     create or replace table {prefix}_rx_encounter as
#     {query_event_rx(
#       code_list_table = code_list_table,
#       rx_version = rx_version, 
#       start_date = query_start_date, end_date = query_end_date
#       )
#     }
#     ;
# ")
# execute_sql(sql_rx)
# toc()
```


```{r}
sql = glue("
    select e.event, e.visit_setting_of_care, e.code, c.code_description, count(*) as n, count(distinct upk_key2) as n_patient
    from {prefix}_mx_encounter e
    left join {prefix}_code_list c
    on e.code = c.code and e.event = c.event
    group by e.event, e.visit_setting_of_care, e.code, c.code_description
    order by e.event, n desc
    ;
")
print(event_freq <- read_query(sql))
sheet_write(event_freq, ss = ss, sheet = 'mx_event_frequency')
```

```{r}
sql = glue("
    select e.event, e.code, c.code_description, count(*) as n, count(distinct upk_key2) as n_patient
    from {prefix}_rx_encounter e
    left join {prefix}_code_list c
    on e.code = c.code and e.event = c.event
    group by e.event, e.code, c.code_description
    order by e.event, n desc
    ;
")
print(event_freq <- read_query(sql))
sheet_write(event_freq, ss = ss, sheet = 'rx_event_frequency')
```

```{r step_3_build_oxbryta_cohort}
sql = glue("  
    create or replace table {prefix}_oxbryta_cohort as
    with query as (
      select upk_key2, claim_date, code, codetype
      from {prefix}_rx_encounter where event = 'oxbryta'
    ),
    add_row_number as(
      select upk_key2, claim_date, code, codetype,
        row_number() over(partition by upk_key2 order by claim_date) as n
      from query
    )
    select upk_key2, claim_date as index_date, code, codetype
    from add_row_number
    where n = 1
    ;
")
execute_sql(sql)

print(read_query(glue("select year(index_date) as year, count(*) as n from {prefix}_oxbryta_cohort group by year order by year")))
print(read_query(glue("select source, code, count(*) as n from {prefix}_oxbryta_cohort group by source, code")))
print(read_query(glue("select min(index_date) from {prefix}_oxbryta_cohort")))
```

```{r step_4_build_control_cohort}
seed = 1234
sql = glue("
  create or replace table {prefix}_control_cohort as
  with query as (
    select upk_key2, claim_date, code, codetype
    from {prefix}_mx_encounter 
    where event = 'scd' and upk_key2 not in (select upk_key2 from {prefix}_oxbryta_cohort)
  ),
  first_claim as(
    select upk_key2, claim_date, code, codetype,
      row_number() over(partition by upk_key2 order by claim_date) as n
    from query
    qualify n = 1
  )
  select c.upk_key2, c.claim_date, 
    greatest(start_date + 365, c.claim_date + 1, to_date('2019-12-01')) as first_index_date,
    e.end_date - 365 as last_index_date,
    last_index_date - first_index_date as length,
    uniform(0::float, 1::float, random({seed})) as rng,
    first_index_date + (length * rng)::int as index_date
  from first_claim c
  inner join ce_{mx_version} e
  on c.upk_key2 = e.upk_key2 and c.claim_date between e.start_date and e.end_date - 365
  where e.end_date - e.start_date >= 730 and length > 0
  ;
")
execute_sql(sql)
```


```{r step_5_combine_cohort}
sql = glue("
  create or replace table {prefix}_combined_cohort as
  with combined as (
    select upk_key2, index_date, 'oxbryta' as cohort
    from {prefix}_oxbryta_cohort
    union
    select upk_key2, index_date, 'control' as cohort
    from {prefix}_control_cohort
  )
  select c.upk_key2, cohort, index_date, 
    count_if(c.index_date - 365 between e.start_date and e.end_date
        and c.index_date + 365 between e.start_date and e.end_date) as elig,
      count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'scd') as n_scd_before
  from combined c
  left join ce_{mx_version} e
  on c.upk_key2 = e.upk_key2 
  left join {prefix}_mx_encounter m
  on c.upk_key2 = m.upk_key2 and m.event = 'scd'
  group by c.upk_key2, cohort, index_date
  ;
")
execute_sql(sql)
```

```{r step_6_final_cohort}
sql = glue("
  create or replace table {prefix}_final_cohort as
  select c.upk_key2, cohort, index_date, 
    count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc') as n_voc_b,
    count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit')) as n_voc_b_ip,
    count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Emergency Room Visit')) as n_voc_b_er,
    count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit', 'Emergency Room Visit')) as n_voc_b_iper,
    count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc') as n_voc_f,
    count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit')) as n_voc_f_ip,
    count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Emergency Room Visit')) as n_voc_f_er,
    count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit', 'Emergency Room Visit')) as n_voc_f_iper
  from {prefix}_combined_cohort c
  left join {prefix}_mx_encounter m
  on c.upk_key2 = m.upk_key2 and m.event = 'voc'
  where elig > 0 and n_scd_before > 0
  group by c.upk_key2, cohort, index_date
  ;
")
# sql = glue("
#   create or replace table {prefix}_final_cohort as
#   select c.upk_key2, cohort, index_date, 
#     count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc') as n_voc_b,
#     count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit')) as n_voc_b_ip,
#     count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Emergency Room Visit')) as n_voc_b_er,
#     count_if(m.claim_date between c.index_date - 365 and c.index_date and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit', 'Emergency Room Visit')) as n_voc_b_iper,
#     count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc') as n_voc_f,
#     count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit')) as n_voc_f_ip,
#     count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Emergency Room Visit')) as n_voc_f_er,
#     count_if(m.claim_date between c.index_date + 1 and c.index_date + 365 and m.event = 'voc' and visit_setting_of_care in ('Inpatient Visit', 'Emergency Room Visit')) as n_voc_f_iper
#   from {prefix}_combined_cohort c
#   left join {prefix}_mx_encounter m
#   on c.upk_key2 = m.upk_key2 and m.event = 'voc'
#   where elig > 0 and n_scd_before > 0
#   group by c.upk_key2, cohort, index_date
#   ;
# ")
execute_sql(sql)
```

```{r step_7_cohort_summary}
sql = glue("
    select
      cohort,
      count(*) as n, 
      count_if(n_voc_b > 0) as voc_1y_before, 
      count_if(n_voc_b_ip > 0) as voc_ip_1y_before, 
      count_if(n_voc_b_er > 0) as voc_er_1y_before, 
      count_if(n_voc_b_iper > 0) as voc_ip_or_er_1y_before, 
      count_if(n_voc_f > 0) as voc_1y_after,
      count_if(n_voc_f_ip > 0) as voc_ip_1y_after,
      count_if(n_voc_f_er > 0) as voc_er_1y_after,
      count_if(n_voc_f_iper > 0) as voc_ip_or_er_1y_after
    from {prefix}_final_cohort
    group by cohort
    ;
")
print(voc_scd_treated <- read_query(sql))
sheet_write(voc_scd_treated, ss = ss, sheet = 'cohort_outcomes')
```


```{r step_7_cohort_summary2}
sql = glue("
    select
      cohort, year(index_date) as year, 
      count(*) as n, 
      count_if(n_voc_b > 0) as voc_1y_before, 
      count_if(n_voc_b_ip > 0) as voc_ip_1y_before, 
      count_if(n_voc_b_er > 0) as voc_er_1y_before, 
      count_if(n_voc_b_iper > 0) as voc_ip_or_er_1y_before, 
      count_if(n_voc_f > 0) as voc_1y_after,
      count_if(n_voc_f_ip > 0) as voc_ip_1y_after,
      count_if(n_voc_f_er > 0) as voc_er_1y_after,
      count_if(n_voc_f_iper > 0) as voc_ip_or_er_1y_after
    from {prefix}_final_cohort
    group by cohort, year
    order by cohort, year
    ;
")
print(voc_scd_treated <- read_query(sql))
sheet_write(voc_scd_treated, ss = ss, sheet = 'cohort_outcomes_by_year')
```

```{r step_7_cohort_summary3}
sql = glue("
    select
      cohort, year(index_date) as year, 
      count(*) as n, 
      count_if(n_voc_b > 0) as voc_1y_before, 
      count_if(n_voc_b_ip > 0) as voc_ip_1y_before, 
      count_if(n_voc_b_er > 0) as voc_er_1y_before, 
      count_if(n_voc_b_iper > 0) as voc_ip_or_er_1y_before, 
      count_if(n_voc_f > 0) as voc_1y_after,
      count_if(n_voc_f_ip > 0) as voc_ip_1y_after,
      count_if(n_voc_f_er > 0) as voc_er_1y_after,
      count_if(n_voc_f_iper > 0) as voc_ip_or_er_1y_after
    from {prefix}_final_cohort
    group by cohort, year
    order by cohort, year
    ;
")
print(voc_scd_treated <- read_query(sql))
sheet_write(voc_scd_treated, ss = ss, sheet = 'cohort_outcomes_by_year')
```
```{r step_8_demo_summary}
sql = glue("
    select
      cohort, year(index_date) as year, 
      count(*) as n, 
      count_if(n_voc_b > 0) as voc_1y_before, 
      count_if(n_voc_b_ip > 0) as voc_ip_1y_before, 
      count_if(n_voc_b_er > 0) as voc_er_1y_before, 
      count_if(n_voc_b_iper > 0) as voc_ip_or_er_1y_before, 
      count_if(n_voc_f > 0) as voc_1y_after,
      count_if(n_voc_f_ip > 0) as voc_ip_1y_after,
      count_if(n_voc_f_er > 0) as voc_er_1y_after,
      count_if(n_voc_f_iper > 0) as voc_ip_or_er_1y_after
    from {prefix}_final_cohort
    group by cohort, year
    order by cohort, year
    ;
")
print(voc_scd_treated <- read_query(sql))
sheet_write(voc_scd_treated, ss = ss, sheet = 'cohort_outcomes_by_year')
```